Large-scale feature selection with Gaussian mixture models for remote sensing images

### 2016-10-14 paper.tex
misc

        \emph{The Bhattacharyya distance} is defined in the case of Gaussian model as
        \begin{align}
            {B}_{cc^\prime} = &\frac{1}{8} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t \left( \frac{\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}}{2} \right)^{-1} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) \nonumber \\
            &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}|}{\sqrt{|\boldsymbol{\Sigma}_c| |\boldsymbol{\Sigma}_{c^\prime}|}} \right).
        \end{align}

@@

        \begin{algorithm}
        \caption{Sequential forward features selection\label{alg:sfs}}
        {\footnotesize
        \begin{algorithmic}[1]
        \REQUIRE $\Omega,J,\text{maxVarNb}$
        \STATE $\Omega=\emptyset$
        \STATE $F=\text{\{all variables\}}$
        \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
        \FORALL{$f_i \in F$}
        \STATE $R_i = J(\{\Omega + f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \STATE $\Omega = \{\Omega + f_j\}$
        \STATE $F = F \setminus f_j$
        \ENDWHILE
        \RETURN $\Omega$
        \end{algorithmic}
        }
        \end{algorithm}

@@

        \begin{algorithm}
        \caption{Sequential floating forward features selection\label{alg:sffs}}
        {\footnotesize
        \begin{algorithmic}[1]
        \REQUIRE $J,\text{maxVarNb}$
        \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
        \STATE $F=\text{\{all variables\}}$
        \STATE $k=0$
        \WHILE{$k \neq \text{maxVarNb}$}
        \FORALL{$f_i \in F$}
        \STATE $R_i = J(\{\Omega_k + f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \STATE $k=k+1$
        \IF{$R_j \geq J(\Omega_k)$}
        \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
        \STATE $\text{flag}=1$
        \WHILE{$k > 2 \text{ and } \text{flag}=1$}
        \FORALL{$f_i \in \Omega_k$}
        \STATE $R_i = J(\{\Omega_k \setminus f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \IF{$R_j > J(\Omega_{k-1})$}
        \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
        \STATE $k=k-1$
        \ELSE
        \STATE $\text{flag}=0$
        \ENDIF
        \ENDWHILE
        \ENDIF
        \ENDWHILE
        \RETURN $\Omega_{\text{maxVarNb}}$
        \end{algorithmic}
        }
        \end{algorithm}

@@

      \begin{prop}[Jeffries-Matusita distance~(\ref{eq:jm})]
      \begin{align}
        \label{eq:jm-update}
            \text{B}_{cc^\prime} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_c \alpha_{c^\prime}}} \right)}_{\substack{\text{computed for each augmented set}}},
      \end{align}
    \end{prop}
    \noindent where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha_c$ and $\mathbf{v}_c$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}_c$.

@@

    \begin{algorithm}
    \caption{Sequential floating forward features selection with updates\label{alg:sffs-update}}
    {\footnotesize
    \begin{algorithmic}[1]
    \REQUIRE $J,\text{maxVarNb}$
    \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
    \STATE $F=\text{\{all variables\}}$
    \STATE $k=0$
    \WHILE{$k \neq \text{maxVarNb}$}
    \FORALL{$c \in \{1,...,C\}$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)}_c = \mathbf{P}_c \boldsymbol{\Lambda}_c \mathbf{P}_c^t$
    \STATE {\bfseries for all} $\lambda_{c}(i)$ {\bfseries do} $\lambda_{c}(i) = \max (\text{EPS\_FLT},\lambda_{c}(i))$
    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)}_c)^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)}_c)$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
    \ENDFOR
    \FORALL{$f_i \in F$}
    \FORALL{$c \in \{1,...,C\}$}
    \STATE Compute update constant $\alpha_c$
    \STATE $\alpha_c = \max (\text{EPS\_FLT},\alpha_c)$
    \ENDFOR
    \STATE $R_i = J(\{\Omega_k + f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \STATE $k=k+1$
    \IF{$R_j \geq J(\Omega_k)$}
    \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
    \STATE $\text{flag}=1$
    \WHILE{$k > 2 \text{ and } \text{flag}=1$}
    \FORALL{$c \in \{1,...,C\}$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)}_c = \mathbf{P}_c \boldsymbol{\Lambda}_c \mathbf{P}_c^t$
    \STATE {\bfseries for all} $\lambda_{c}(i)$ {\bfseries do} $\lambda_{c}(i) = \max (\text{EPS\_FLT},\lambda_{c}(i))$
    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)}_c)^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)}_c)$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
    \ENDFOR
    \FORALL{$f_i \in \Omega_k$}
    \FORALL{$c \in \{1,...,C\}$}
    \STATE Compute update constant $\alpha_c$
    \STATE $\alpha_c = \max (\text{EPS\_FLT},\alpha_c)$
    \ENDFOR
    \STATE $R_i = J(\{\Omega_k \setminus f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \IF{$R_j > J(\Omega_{k-1})$}
    \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
    \STATE $k=k-1$
    \ELSE
    \STATE $\text{flag}=0$
    \ENDIF
    \ENDWHILE
    \ENDIF
    \ENDWHILE
    \RETURN $\Omega_{\text{maxVarNb}}$
    \end{algorithmic}
    }
    \end{algorithm}

@@

    When creating training and validation  sets, special care is taken
    to assure that training samples are picked out from distinct areas
    than test  samples.  The  polygons of the  reference are  split in
    smaller polygons and then 50\%  of the polygons are taken randomly
    for training and the remaining 50\% for validation.  An example of
    training      and     validation      set     is      shown     in
    Figure~\ref{fig:set-aisa}.  From the  training  polygons, a  given
    number of samples  were selected to build the  training set, while
    all  the pixels  from the  validation polygons  were used  for the
    validation.  Moreover 20  random trials were run  with a different
    training  set (different  polygons).  Table~\ref{tab:aisa-otbsimu}
    presents  the results  of the  experiment with  mean and  standard
    deviation  of  the  Kappa  coefficient  over  the  20  trials  and
    Table~\ref{tab:aisa-otbsimu-time}  the   corresponding  processing
    time. Additionally, Figure~\ref{fig:meanNbVar-aisa} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.

@@

            GMM SFS kappa & {\bfseries 0.684 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.711 (0.028)} \\
            GMM SFS JM &    {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
            GMM SFFS JM &   {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
            GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
            KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
            Random Forest & 0.645 (0.026) & 0.673 (0.023) & {\bfseries 0.693 (0.023)} \\
            \bottomrule
        \end{tabular}
    \end{table}

@@

    For the Potsdam  dataset, training samples were  selected from one
    tile (5\_11)  and validation  samples were all  the pixel  of tile
    5\_12     or    3\_10.      Table~\ref{tab:potsdam-otbsimu}    and
    Table~\ref{tab:potsdam-otbsimu-big} present  the results  in terms
    of classification accuracy and processing time.

@@

\begin{proof}[Proof of proposition~(\ref{eq:update-quad})]
    {\footnotesize
    \begin{alignat*}{3}
    (&\mathbf{x}^{(k)})^t && (\boldsymbol{\Sigma}^{(k)}_c)^{-1} \mathbf{x}^{(k)}
       = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
        \left[\begin{array}{cc}
            \mathbf{A}_c   & \mathbf{v}_c \\
            \mathbf{v}_c^t & \frac{1}{\alpha_c}
        \end{array}\right]
        \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t \mathbf{A}_c + x_k \mathbf{v}_c^t & (\mathbf{x}^{(k-1)})^t \mathbf{v}_c + \frac{x_k}{\alpha_c} \end{array}\right]
            \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
     &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A}_c \mathbf{x}^{(k-1)} + x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + (\mathbf{x}^{(k-1)})^t \mathbf{v}_c x_k + \frac{(x_k)^2}{\alpha_c} \\
     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} + \frac{1}{\alpha_c} (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
     & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} + \alpha_c \mathbf{v}_c \mathbf{v}_c^t) \mathbf{x}^{(k-1)}\\
     & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c \mathbf{v}_c^t \mathbf{x}^{(k-1)} \\
     & &&+ 2 \frac{x_k}{\alpha_c} \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c^2}) \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c + \frac{x_k}{\alpha_c} )^2 \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] \mathbf{x}^{(k)} )^2
    \end{alignat*}
    }
\end{proof}

>>>>

        \emph{The Bhattacharyya distance} is defined in the case of Gaussian model as
        \begin{align}
            {B}_{cc^\prime} = &\frac{1}{8} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t \left( \frac{\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}}{2} \right)^{-1} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) \nonumber \\
            &+ \frac{1}{2} \log \left( \frac{|\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}|}{\sqrt{|\boldsymbol{\Sigma}_c| |\boldsymbol{\Sigma}_{c^\prime}|}} \right).
        \end{align}

@@

        \begin{algorithm}
        \caption{Sequential forward features selection\label{alg:sfs}}
        {\footnotesize
        \begin{algorithmic}[1]
        \REQUIRE $\Omega,J,\text{maxVarNb}$
        \STATE $\Omega=\emptyset$
        \STATE $F=\text{\{all variables $f_i$\}}$
        \WHILE{$\text{card}(\Omega) < maxVarNb$}
        \FORALL{$f_i \in F$}
        \STATE $R_i = J(\{\Omega + f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \STATE $\Omega = \{\Omega + f_j\}$
        \STATE $F = F \setminus f_j$
        \ENDWHILE
        \RETURN $\Omega$
        \end{algorithmic}
        }
        \end{algorithm}

@@

        \begin{algorithm}
        \caption{Sequential floating forward features selection\label{alg:sffs}}
        {\footnotesize
        \begin{algorithmic}[1]
        \REQUIRE $J,\text{maxVarNb}$
        \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
        \STATE $F=\text{\{all variables $f_i$\}}$
        \STATE $k=0$
        \WHILE{$k < \text{maxVarNb}$}
        \FORALL{$f_i \in F$}
        \STATE $R_i = J(\{\Omega_k + f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \STATE $k=k+1$
        \IF{$R_j \geq J(\Omega_k)$}
        \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
        \STATE $\text{flag}=1$
        \WHILE{$k > 2 \text{ and } \text{flag}=1$}
        \FORALL{$f_i \in \Omega_k$}
        \STATE $R_i = J(\{\Omega_k \setminus f_i\})$
        \ENDFOR
        \STATE $j=\text{arg} \max_{i} R_i$
        \IF{$R_j > J(\Omega_{k-1})$}
        \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
        \STATE $k=k-1$
        \ELSE
        \STATE $\text{flag}=0$
        \ENDIF
        \ENDWHILE
        \ENDIF
        \ENDWHILE
        \RETURN $\Omega_{\text{maxVarNb}}$
        \end{algorithmic}
        }
        \end{algorithm}

@@

      \begin{prop}[Jeffries-Matusita distance~(\ref{eq:jm})]
      \begin{align}
        \label{eq:jm-update}
            \text{B}_{cc^\prime} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{2} \log \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 + \frac{1}{2} \log \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_c \alpha_{c^\prime}}} \right)}_{\substack{\text{computed for each augmented set}}},
      \end{align}
    \end{prop}
    \noindent where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha_c$ and $\mathbf{v}_c$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}_c$.

@@

    \begin{algorithm}
    \caption{Sequential floating forward features selection with updates\label{alg:sffs-update}}
    {\footnotesize
    \begin{algorithmic}[1]
    \REQUIRE $J,\text{maxVarNb}$
    \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
    \STATE $F=\text{\{all variables $f_i$\}}$
    \STATE $k=0$
    \WHILE{$k < \text{maxVarNb}$}
    \FORALL{$c \in \{1,...,C\}$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)}_c = \mathbf{P}_c \boldsymbol{\Lambda}_c \mathbf{P}_c^t$
    \STATE {\bfseries for all} $\lambda_{c}(i)$ {\bfseries do} $\lambda_{c}(i) = \max (\text{EPS\_FLT},\lambda_{c}(i))$
    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)}_c)^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)}_c)$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
    \ENDFOR
    \FORALL{$f_i \in F$}
    \FORALL{$c \in \{1,...,C\}$}
    \STATE Compute update constant $\alpha_c$
    \STATE $\alpha_c = \max (\text{EPS\_FLT},\alpha_c)$
    \ENDFOR
    \STATE $R_i = J(\{\Omega_k + f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \STATE $k=k+1$
    \IF{$R_j \geq J(\Omega_k)$}
    \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
    \STATE $\text{flag}=1$
    \WHILE{$k > 2 \text{ and } \text{flag}=1$}
    \FORALL{$c \in \{1,...,C\}$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)}_c = \mathbf{P}_c \boldsymbol{\Lambda}_c \mathbf{P}_c^t$
    \STATE {\bfseries for all} $\lambda_{c}(i)$ {\bfseries do} $\lambda_{c}(i) = \max (\text{EPS\_FLT},\lambda_{c}(i))$
    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)}_c)^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)}_c)$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right)$} using Propositions (\ref{eq:update-inv}), (\ref{eq:update-quad}) and (\ref{eq:update-log})
    \ENDFOR
    \FORALL{$f_i \in \Omega_k$}
    \FORALL{$c \in \{1,...,C\}$}
    \STATE Compute update constant $\alpha_c$
    \STATE $\alpha_c = \max (\text{EPS\_FLT},\alpha_c)$
    \ENDFOR
    \STATE $R_i = J(\{\Omega_k \setminus f_i\})$ using Equations (\ref{eq:q-update}), (\ref{eq:skl-update}) or (\ref{eq:jm-update})
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \IF{$R_j > J(\Omega_{k-1})$}
    \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
    \STATE $k=k-1$
    \ELSE
    \STATE $\text{flag}=0$
    \ENDIF
    \ENDWHILE
    \ENDIF
    \ENDWHILE
    \RETURN $\Omega_{\text{maxVarNb}}$
    \end{algorithmic}
    }
    \end{algorithm}

@@

    When creating training and validation  sets, special care is taken
    to assure that training samples are picked out from distinct areas
    than test  samples.  The  polygons of the  reference are  split in
    smaller polygons and then 50\%  of the polygons are taken randomly
    for training and the remaining 50\% for validation.  An example of
    training      and     validation      set     is      shown     in
    Figure~\ref{fig:set-aisa}.  From the  training  polygons, a  given
    number of samples  were selected to build the  training set, while
    all  the pixels  from the  validation polygons  were used  for the
    validation.  Moreover 20  random trials were run  with a different
    training  set (different  polygons).  Table~\ref{tab:aisa-otbsimu}
    presents  the results  of the  experiment with  mean and  standard
    deviation  of  the  Kappa  coefficient  over  the  20  trials  and
    Table~\ref{tab:aisa-otbsimu-time}  the   corresponding  processing
    time. Bold values corresponds to best results. In Table~\ref{tab:aisa-otbsimu}, when several bold scores appears for the same experiment, it means that the scores has been assessed as equivalent with a Wilcoxon rank-sum test \cite{mann1947test}. Additionally, Figure~\ref{fig:meanNbVar-aisa} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.

@@

            GMM SFS kappa & {\bfseries 0.678 (0.029)} & {\bfseries 0.687 (0.029)} & {\bfseries 0.699 (0.028)} \\
            GMM SFS JM &    {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.701 (0.029)} \\
            GMM SFFS JM &   {\bfseries 0.685 (0.030)} & {\bfseries 0.689 (0.030)} & {\bfseries 0.701 (0.029)} \\
            GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
            KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
            Random Forest & 0.645 (0.026) & 0.673 (0.023) & {\bfseries 0.693 (0.023)} \\
            \bottomrule
        \end{tabular}
    \end{table}

    % this one is result without changing the number of selected variables
    % \begin{table}[!t]
    %     \centering
    %     \caption{Average classification accuracy 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
    %     \begin{tabular}{lccc}\toprule
    %          & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-4}
    %         \# samples by class & 250 & 500 & 1000 \\ \midrule

    %         GMM SFS kappa & {\bfseries 0.684 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.711 (0.028)} \\
    %         GMM SFS JM &    {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
    %         GMM SFFS JM &   {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
    %         GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
    %         KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
    %         Random Forest & 0.645 (0.026) & 0.673 (0.023) & 0.693 (0.023) \\
    %         \bottomrule
    %     \end{tabular}
    % \end{table}

@@

    For the Potsdam  dataset, training samples were  selected from one
    tile (5\_11)  and validation  samples were all  the pixel  of tile
    5\_12     or    3\_10.      Table~\ref{tab:potsdam-otbsimu}    and
    Table~\ref{tab:potsdam-otbsimu-big} present  the results  in terms
    of classification accuracy and processing time. Bold values corresponds to best results. In Table~\ref{tab:potsdam-otbsimu}, when several bold scores appears for the same experiment, it means that the scores has been assessed as equivalent with a Wilcoxon rank-sum test \cite{mann1947test}.

@@

\begin{proof}[Proof of proposition~(\ref{eq:update-quad})]
    \begin{alignat*}{3}
    (&\mathbf{x}^{(k)})^t && (\boldsymbol{\Sigma}^{(k)}_c)^{-1} \mathbf{x}^{(k)} \\
     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
        \left[\begin{array}{cc}
            \mathbf{A}_c   & \mathbf{v}_c \\
            \mathbf{v}_c^t & \frac{1}{\alpha_c}
        \end{array}\right]
        \left[\begin{array}{c} \mathbf{x}^{(k-1)} \\ x_k \end{array}\right] \\
     &= &&\left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x_k \end{array}\right]
            \left[\begin{array}{c} \mathbf{A}_c \mathbf{x}^{(k-1)} + x_k \mathbf{v}_c \\ \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{x_k}{\alpha_c} \end{array}\right] \\
     &= &&(\mathbf{x}^{(k-1)})^t \mathbf{A}_c \mathbf{x}^{(k-1)} + x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} \\
     & &&+ (\mathbf{x}^{(k-1)})^t \mathbf{v}_c x_k + \frac{(x_k)^2}{\alpha_c} \\
     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \\
     & &&+ \frac{1}{\alpha_c} (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}) \mathbf{x}^{(k-1)}\\
     & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
     &= &&(\mathbf{x}^{(k-1)})^t ((\boldsymbol{\Sigma}_c^{(k-1)})^{-1} + \alpha_c \mathbf{v}_c \mathbf{v}_c^t) \mathbf{x}^{(k-1)}\\
     & &&+ 2 x_k \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c} \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c \mathbf{v}_c^t \mathbf{x}^{(k-1)} \\
     & &&+ 2 \frac{x_k}{\alpha_c} \mathbf{v}_c^t \mathbf{x}^{(k-1)} + \frac{(x_k)^2}{\alpha_c^2}) \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( (\mathbf{x}^{(k-1)})^t \mathbf{v}_c + \frac{x_k}{\alpha_c} )^2 \\
     &= &&(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{x}^{(k-1)} + \alpha_c ( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] \mathbf{x}^{(k)} )^2
    \end{alignat*}
\end{proof}
